Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-25 TACAS2018/paper.tex
description of properties

Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer,Franklin15}.  Correct synthesis is non-trivial, even in cases with
restricted dynamics.  We examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.  However,
the use of digital control architectures adds new challenges caused by
artifacts specific to digital control, such as the effects of
finite-precision arithmetic, time discretization and quantization errors
introduced by A/D and D/A conversion.

@@

\noindent\textbf{Discretization Effects --}
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and finite word length (FWL) semantics, with
the goal either to verify (e.g.,~\cite{daes20161}) or to optimize
(e.g.,~\cite{oudjida2014design}) given implementations.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big\{ \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \, \forall k\in [0,n-1], \,
\vec{x}_0 \in X_0, \,
\vec{u}_{k} \in U  \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations (transitive closure \note{the T.Closure refers to the fixpoint, right? That is, it is what is defined next.}). 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
generalises the previous notions over an unbounded time horizon.

@@

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The feedback controller is driven by the observer state $\hat{\vec{x}}$, 
and is characterised by a gain matrix $\mat{K}$ \note{has index in figure} that influences the closed-loop dynamics.  
Moreover, $\vec{y}_e=\hat{\vec{y}}_k - \vec{y}_k$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
  whereas $\mat{L}$ \note{with index in figure} is an error gain matrix that amplifies the output estimation error. 
Figure~\ref{fig:observersystem} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}$ and $\mat{L}$.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the system at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}$ and $\mat{L}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time. 
 In order to do so, 
 we accelerate the dynamics of the system according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x} \text{ with } k \in \mathbb{R}^+$ \note{naturals N?} is a witness of the system at time $t=kT_d$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \mat{B}\vec{u}(t)=\vec{u}(0)$. 
 For such models, a witness $\vec{x}_T$ is computed next.  
 \note{Need to assume A invertible?}
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} \text{ where } \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 is a witness of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) \text{ with } k \in \mathbb{R}^+$ \note{naturals N?} is the state of the system at time $kT_d$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time, 
namely a new input value is produced only at every sample time but implemented onto the continuous-time plant dynamics. 
 \note{Need to assume A invertible?}
%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 and where $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ and 
 $\mat{A}_d, \mat{B}_d$ describe the discretized plant dynamics, 
 is the state of the model 
 \begin{align}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align}
 \end{theorem}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

In our model delays are all deferred to the ZOH \note{acronym undefined - define these and others, with reference to the Figure 1 above}, 
i.e., we presume that \note{qualitative: everything that happens in the digital world takes place at
the sampling instance $t$} and the output of the result is delayed at the DAC 
when it is fed back into the real world.  
We will also presume that this delay is much smaller than the sampling time \note{does this assumption make practical sense? why? add reference?}.

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upper bound by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm (\note{what? unclear?} in for sets, $\norm{X}_1 = sup \{\norm{\vec{x}}_1 \mid \vec{x} \in X\}$). 
\end{theorem}

\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
We over-approximate the effects of this vector by introducing a noise signal at the output $\nu_t=[-q_t, q_t]$ and use the dynamics of the system without delay.

@@

In Figure \ref{fig:observersystem2} we introduced several sources of noise,
which are due to quantization effects.  These effects are caused by the use
of imprecise numerical representations over the reals in the digital domain
of the controller which is an important consideration when trying to synthesise
sound controllers.
The models we consider have two sources of error that are due to numerical  
representation which we commonly refer to as FWL \note{undefined acronym} effects.
The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

@@

The ability of a system to remain within a bounded set of states is called stability.
In this paper, we are interested in
%% There are different types of stability, such as Bounded Input/Bounded Output (BIBO)
%% in which the requirement is that given a bounded input the output remains within
%% certain limits, or
asymptothic stability, where the output is expected to
converge to a certain value. %% In our model, we are concerned with the latter
%% type of stability is directly related to
%% convergence, which means it directly relates to the eigenvalues of the dynamics.
As we want to use Jury's criterion~\cite{fadali}\footnote{More details on Jury's criterion can be found in~\ref{sec:closed_stability}}, which is expressed 
by an easy algebraic formula, we must ... \note{summarise what we do in the rest of this section.}

@@

\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
%
We now use this limit to express a constraint in terms of the
coefficients of the characteristic polynomial.  Let this
polynomial be:
%
\begin{equation}
P(s)=\prod_{i=0}^p (s-\lambda_i)=\sum_{j=0}^p c_js^j.
\end{equation}
%
%We can use Jury's criterion to limit $|\lambda_i| \leq 1$.
Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
to $P'(s)=\prod_{i=0}^p (s-\lambda_i')=\sum_{j=0}^p c_j's^j$ where $c_j'$ are
unknown.  By replacing $\lambda_i$, we obtain
%
\begin{equation}
P'=\prod_{i=0}^p (s-\lambda_i')=\prod_{i=0}^p \left(s-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^p \frac{c_j}{\overline{\lambda}^{p-j}}s^j
\end{equation}
%
where all members are known. Thus, we can ensure that the settling time is smaller than
a certain magnitude by ensuring that 
the eigenvalues are smaller too (which we can do with Jury's criterion).

@@

Our approach is illustrated in Figure~\ref{fig:CEGIS-OR} and consists of four main
phases:
\begin{itemize}
\item {\sc Abstract}: constructs an unsound abstract model that over-approximates
the dynamics of the system.
\note{what does it mean to be over-approximating but unsound?}
\item {\sc synthesize}: generates a candidate controller and observer.
  As it makes use of the potentially coarse abstraction constructed by {\sc Abstract},
  the candidate doesn't necessarily meet the stability or safety specification.
  %% these do not necessarily meet any portion
  %% of the specification, but either through optimization or refinement become
  %% more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines the abstraction needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
  \note{{\sc abstract} doesn't need an abstraction.}
  \note{how about optimization?}
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon. If the verification fails, produces a counterexample
  describing the reason for failure.
\end{itemize}

%% The first one aims at making the system as
%% convergent as possible.  This does not only ensure stability, but reduces
%% the response time of the system, moving away from the unsafe states faster.
%% The second one seeks to minimize rotations.  If the system rotates slower, it
%% is less likely to reach unsafe states before converging.
%% In both cases, the shape of the eigenspace may affect us negativey, so we
%% use a measure of the size of the reach space to guide our optimization.

%The verifier's abstraction uses a CEGAR loop in order to minimise 
%over-approximation errors and provides counterexamples for both
%discrete and continuous dynamics to feed a CEGIS loop on the synthesiser
%waiting for a new candidate solution. \note{not sure about this sentence.
%  Which is the CEGAR loop?}

@@

\note{In order to show the improvements due to our optimization-enhanced CEGIS, we also compared our approach against a standard CEGIS procedure..}

\note{remove some of the subcases of CEGIS and CEGIS-OR. Those that are still used need to be explained in the actual section, not in the table's caption.}

@@

For the majority of our benchmarks, we observe that our approach is
significantly faster than using standard CEGIS. We further note the cases where
standard CEGIS times out, which represent an improvement of several orders of
magnitude for our optimization approach.  The magnitude of the improvement is 
largely dependent on the shape of the problem. 
%When the tighter restriction isdependent on the optimization parameters the overall speed is

>>>>

Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer,Franklin15}.  Correct synthesis is non-trivial, even in cases with
restricted dynamics.  We examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.  However,
the use of digital control architectures adds new challenges caused by
artifacts specific to digital control, such as the effects of
finite-precision arithmetic (referred to here as Finite Word Length or FWL effects), time discretization and quantization errors
introduced by A/D and D/A conversion.

@@

\noindent\textbf{Discretization Effects --}
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and Finite Word Length (FWL) semantics, with
the goal either to verify (e.g.,~\cite{daes20161}) or to optimize
(e.g.,~\cite{oudjida2014design}) given implementations.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big\{ \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \, \forall k\in [0,n-1], \,
\vec{x}_0 \in X_0, \,
\vec{u}_{k} \in U  \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
generalises the previous notions over an unbounded time horizon (transitive closure).

@@

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The feedback controller is driven by the observer state $\hat{\vec{x}}$, 
and is characterised by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics.  
Moreover, $\vec{y}_e=\hat{\vec{y}}_k - \vec{y}_k$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
  whereas $\mat{L}_{\langle C \rangle}$ is an error gain matrix that amplifies the output estimation error. 
Figure~\ref{fig:observersystem} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the system at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the system according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x} \text{ with } T \in \mathbb{R}^+$ is the state of the system at time $t=T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \mat{B}\vec{u}(t)=\vec{u}(0)$. 
 We will assume $\mat{A}$ is invertible when there are geometric multiplicities $>1$.
 For such models, a the state $\vec{x}_T$ is computed next.  
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} \text{ where } \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 is the state of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}\vec{u}) \text{ with } T \in \mathbb{R}^+$ is the state of the system at time $T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time, 
namely a new input value is produced only at every sample time but implemented onto the continuous-time plant dynamics. 
We will assume $\mat{A}$ is invertible, since otherwise $B_T$ would have to be calculated by integration which is impractical.

%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 and where $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ and 
 $\mat{A}_d, \mat{B}_d$ describe the discretized plant dynamics, 
 is the state of the model 
 \begin{align}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align}
 \end{theorem}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

In our model, delays are all deferred to the Zero Order Holder (ZOH) of the DAC (See Figure \ref{fig:observersystem2}). 
Everything that happens in the digital world takes place at times $t=kT_s, k \in \mathbb{N}$ (discrete time progression),
and the output of the result is delayed at the DAC when it is fed back into the continuous time domain at time ($t=kT_s+t_d$).

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upper bound by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm (which we extend to sets as $\norm{X}_1 = sup \{\norm{\vec{x}}_1 \mid \vec{x} \in X\}$). 
\end{theorem}

\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
We over-approximate the effects of this vector by introducing a noise signal at the output $\nu_t=[-q_t, q_t]$ and use the dynamics of the system without delay.  This noise will be small when the delay is much smaller than the sampling time.

@@

In Figure \ref{fig:observersystem2} we introduced several sources of noise,
which are due to quantization effects.  These effects are caused by the use
of imprecise numerical representations over the reals in the digital domain
of the controller which is an important consideration when trying to synthesise
sound controllers.
The models we consider have two sources of error that are due to Finite Word Length (FWL)
representation.
The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

@@

The ability of a system to remain within a bounded set of states is called stability.
In this paper, we are interested in asymptothic stability, where the output is expected to
converge to a certain value.  
As we want to use Jury's criterion~\cite{fadali}\footnote{More details on Jury's criterion can be found in~\ref{sec:closed_stability}}, which is expressed 
by an easy algebraic formula, we transform the model into a representation that contains the desired coefficients expressed as sums of values (no multiplications).  Such representations are called Controllable and Observable Canonical Forms~\cite{Astrom08} \footnote{See~\ref{sec:reachable} and~\ref{sec:observable}}.  In this section we explore the implications of using these canonical forms alongside an FWL representation.

@@

\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
%
We now use this limit to express a constraint in terms of the
coefficients of the characteristic polynomial.  Let this
polynomial be:
%
\begin{equation}
P(s)=\prod_{i=0}^p (s-\lambda_i)=\sum_{j=0}^p c_js^j.
\end{equation}
%
%We can use Jury's criterion to limit $|\lambda_i| \leq 1$.
Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
to $P'(s)=\prod_{i=0}^p (s-\lambda_i')=\sum_{j=0}^p c_j's^j$ where $c_j'$ are
unknown.  By replacing $\lambda_i$, we obtain
%
\begin{equation}
P'=\prod_{i=0}^p (s-\lambda_i')=\prod_{i=0}^p \left(s-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^p \frac{c_j}{\overline{\lambda}^{p-j}}s^j
\end{equation}
%
where all members are known. Thus, we can ensure that the settling time is smaller than
a certain magnitude by ensuring that 
the eigenvalues are smaller too (which we can do with Jury's criterion).
We will use $\phi_{stability}$ to refer to the combined stability and transient response specifications.

@@

Our approach is illustrated in Figure~\ref{fig:CEGIS-OR} and consists of four main
phases:
\begin{itemize}
\item {\sc Abstract}: constructs an unsound abstract model that approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer.
  As it makes use of the potentially coarse abstraction constructed by {\sc Abstract},
  the candidate doesn't necessarily meet the stability or safety specification.
\item {\sc model}: defines the properties needed by {\sc Abstract} and {\sc Optimise} to create 
  adequate abstractions at each stage in order to reduce the probability
  of failure in the next iteration.  Our optimization procedure aims at reducing the
  size of the reach space of the closed loop system while preserving the verification
  properties (safety, stability and response time).  This phase will first attempt this objective
  and failing this feed the {\sc Abstract} phase the necessary information to produce a better candidate.
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon. If the verification fails, produces a counterexample describing the reason for failure.
\end{itemize}

@@

@@

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority of
our benchmarks, we observe that our approach is significantly faster than
standard CEGIS. We further note the cases where standard CEGIS times out, which
represent an improvement of several orders of magnitude for our optimization
approach.  The magnitude of the improvement is largely dependent on the shape
of the problem.
